---
title: "Pika-Yak Interaction: Plant Cover"
author: "Douglas Lawton"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-cap-location: bottom
    theme: cosmo
editor: source
---

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(mgcv)
library(knitr)
library(kableExtra)

# --- Set Root Directory for Project ---
here::i_am("README.md")  # Anchor project for reproducible paths

# --- Custom Color Palette ---
pink <- '#ff8080'
green <- '#2fc09b'

```

```{r}
#| echo: false
#| label: read in data
#| results: "hide"
#| message: false
#| warning: false
cover_data <- read_csv(here('data/processed/experiment_plant_cover/plant_cover_by_treatment.csv'))

forbs_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/emms.csv'))
forbs_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/posthoc_letters.csv'))
forbs_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/model_summary.csv'))
forbs_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/posthoc_contrasts.csv'))

grasses_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/emms.csv'))
grasses_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/posthoc_letters.csv'))
grasses_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/model_summary.csv'))
grasses_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/posthoc_contrasts.csv'))

poison_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/emms.csv'))
poison_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/posthoc_letters.csv'))
poison_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/model_summary.csv'))
poison__model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/posthoc_contrasts.csv'))

sedges_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/emms.csv'))
sedges_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/posthoc_letters.csv'))
sedges_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/model_summary.csv'))
sedges_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/posthoc_contrasts.csv'))

```

```{r}
#| label: fig-plant-cover
#| fig-cap: "Plant cover between experimental plots"
#| fig-width: 8
#| fig-height: 8
#| echo: false
#| reulsts: asis
#| message: false
#| warning: false

make_cover_plot <- function(plant_name, letters_df, y_label, text_y = 41, letters = TRUE) {
  p <- cover_data |>
    filter(plant == plant_name) |>
    left_join(letters_df, by = c("pika_treatment", "posion_plant_treatment")) |>
    ggplot(aes(x = pika_treatment, y = cover, color = posion_plant_treatment)) +
    geom_jitter(
      size = 1, alpha = 0.4,
      position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7)
    ) +
    geom_point(
        data = letters_df,
        aes(x = pika_treatment, y = response * 100, fill = posion_plant_treatment),
        position = position_dodge(width = 0.7),
        size = 8, pch = 21, color = "black"
    ) +
    scale_fill_manual(values = c(green, pink)) +
    scale_color_manual(values = c(green, pink)) +
    labs(
      y = y_label,
      x = "",
      color = "Poison plant treatment",
      fill = "Poison plant treatment"
    ) +
    theme_pubr(base_size = 15) +
    theme(legend.position = "bottom", legend.title = element_blank())

  if (letters) {
    p <- p  +
      geom_text(
        data = letters_df,
        aes(x = pika_treatment, y = text_y, label = .group, group = posion_plant_treatment),
        position = position_dodge(width = 0.7),
        size = 5, color = "black", show.legend = FALSE
      )
  }

  p
}


# Now call the function with appropriate arguments
forb_plot    <- make_cover_plot("forbs", forbs_letters, "forb cover (%)", text_y = 41,letters=FALSE)
grass_plot   <- make_cover_plot("grasses", grasses_letters, "grass cover (%)", text_y = 41)
poison_plant <- make_cover_plot("s_chamaejasme", poison_letters, "S. chamaejasme cover (%)", text_y = 43)
sedge_plot   <- make_cover_plot("sedges", sedges_letters, "sedge cover (%)", text_y = 70,letters=FALSE)



(forb_plot + grass_plot) / (poison_plant + sedge_plot) +
    plot_annotation(tag_levels = 'A') + 
    plot_layout(guides = "collect") & theme(legend.position = 'bottom')



```

# Forbs Cover

Forb cover showed a small but significant reduction with the poison plant (*S. chamaejasme*) treatment, while pika presence alone had no measurable effect (@tbl-forbs-model-summary). The interaction between pika and poison plant treatments was not significant, indicating that pika did not alter the poison plant’s effect on forb cover. Random variation across blocks and years was small, and no month-to-month variation was detected.

# Grasses Cover

Grass cover was significantly reduced by the poison plant (*S. chamaejasme*) treatment, with odds of occurrence dropping to less than half in treated plots. Pika presence alone had no detectable effect in untreated plots, but in treated plots, pika presence substantially increased grass cover, offsetting some of the negative effect of the poison plant. This positive interaction suggests that pika may facilitate grass recovery when *S. chamaejasme* is present. Random variation across blocks and months was small, and no variation was detected across years. See the model summary and contrast tables (@tbl-grasses-model-summary, @tbl-grasses-model-contrasts).

# *S. chamaejasme* Cover

*Stellera chamaejasme* cover was strongly increased in the poison plant treatment (@tbl-poison-model-summary). Pika presence alone had no significant effect in untreated plots but significantly reduced the large cover gains from the poison plant treatment (@tbl-poison-model-contrasts). This interaction indicates that pika substantially suppress *S. chamaejasme* when it is present in high abundance. Random variation across blocks, years, and months was small relative to the fixed effects.

# Sedges Cover

Sedge cover was significantly reduced by the poison plant (*S. chamaejasme*) treatment, while pika presence alone had no significant effect. The positive pika × poison plant interaction was marginally non-significant (p = 0.076), suggesting a possible tendency for pika to lessen the negative effect of the poison plant (@tbl-sedges-model-summary). Random variation was small across blocks and months, and no year-to-year variation was detected.


```{r}
#| label: tbl-forbs-model-summary
#| tbl-cap: "Model summary for forb cover."
#| echo: false
#| message: false
#| warning: false

library(kableExtra)
forbs_model_summary |>
  mutate(
    term = case_when(
        term == 'pika_treatmentpika' ~ 'pika',
        term == 'posion_plant_treatmentS. chamaejasme' ~ 'poison plant',
        term == 'pika_treatmentpika:posion_plant_treatmentS. chamaejasme' ~ 'pika : poison treamtent',
        effect == 'ran_pars' ~ group,
        TRUE ~ term
    )
  ) |>
  select(!c(component,group)) |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-forbs-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for forb cover by treatment group."
#| echo: false
#| message: false
#| warning: false
#| results: "hide"

forbs_model_contrasts |>
  kable(format = "html", booktabs = TRUE, digits = 3)|>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-grasses-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false

grasses_model_summary |>
  mutate(
    term = case_when(
        term == 'pika_treatmentpika' ~ 'pika',
        term == 'posion_plant_treatmentS. chamaejasme' ~ 'poison plant',
        term == 'pika_treatmentpika:posion_plant_treatmentS. chamaejasme' ~ 'pika : poison treamtent',
        effect == 'ran_pars' ~ group,
        TRUE ~ term
    )
  ) |>
  select(!c(component,group)) |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-grasses-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

grasses_model_contrasts |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-poison-model-summary
#| tbl-cap: "Model summary for *S. chamaejasme* cover."
#| echo: false
#| message: false
#| warning: false

poison_model_summary |>
  mutate(
    term = case_when(
        term == 'pika_treatmentpika' ~ 'pika',
        term == 'posion_plant_treatmentS. chamaejasme' ~ 'poison plant',
        term == 'pika_treatmentpika:posion_plant_treatmentS. chamaejasme' ~ 'pika : poison treamtent',
        effect == 'ran_pars' ~ group,
        TRUE ~ term
    )
  ) |>
  select(!c(component,group)) |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-poison-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for *S. chamaejasme* cover by treatment group."
#| echo: false
#| message: false
#| warning: false

poison__model_contrasts |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-sedges-model-summary
#| tbl-cap: "Model summary for sedge cover."
#| echo: false
#| message: false
#| warning: false

sedges_model_summary |>
  mutate(
    term = case_when(
        term == 'pika_treatmentpika' ~ 'pika',
        term == 'posion_plant_treatmentS. chamaejasme' ~ 'poison plant',
        term == 'pika_treatmentpika:posion_plant_treatmentS. chamaejasme' ~ 'pika : poison treamtent',
        effect == 'ran_pars' ~ group,
        TRUE ~ term
    )
  ) |>
  select(!c(component,group)) |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```

```{r}
#| label: tbl-sedges-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for sedge cover by treatment group."
#| echo: false
#| message: false
#| warning: false
#| results: "hide"

sedges_model_contrasts |>
  kable(format = "html", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped","scale_down"))
```