---
title: "Methods and Results"
author: "Douglas Lawton"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    fig-cap-location: bottom
    theme: cosmo
  docx:
    toc: false
    number-sections: true
editor: source
crossref:
  custom:
    - kind: float
      key: fig
      latex-env: fig
      reference-prefix: Figure
      space-before-numbering: true
      latex-list-of-description: Figure
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Supplementary Figure 
      space-before-numbering: true
      latex-list-of-description: Supplementary Figure
    - kind: float
      key: tbl
      latex-env: tbl
      reference-prefix: Table
      space-before-numbering: true
      latex-list-of-description: Table
    - kind: float
      key: supptbl
      latex-env: supptbl
      reference-prefix: Supplementary Table
      space-before-numbering: true
      latex-list-of-description: Supplementary Table
---

```{r}
#| label: gloabl settings
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(mgcv)
library(gratia)
library(qs2)
library(knitr)
library(kableExtra)
library(cowplot)
library(glue)
library(gt)


# --- Set Root Directory for Project ---
here::i_am("README.md")  # Anchor project for reproducible paths

# --- Custom Color Palette ---
pink <- '#ff8080'
green <- '#2fc09b'


plant_colors <- c(green, pink)

base_size = 16
geom_text_size = 7
```

```{r reading in data}
#| echo: false
#| warning: false
#| message: false
#| results: "hide"

# ============================================================
#                 DATA LOADING HELPERS
# ============================================================

# Load a CSV relative to project root
load_csv <- function(...) {
  read_csv(here(...), show_col_types = FALSE)
}

# Generalized function for loading model outputs (emms, letters, summary, contrasts)
# Handles optional files (e.g., GAM models don't produce emmeans)
load_model_data <- function(base_dir, name = NULL) {
  base_path <- if (is.null(name)) here(base_dir) else here(glue("{base_dir}/{name}"))
  
  # Helper to safely load CSV if it exists, NULL otherwise
  safe_read <- function(filename) {
    filepath <- file.path(base_path, filename)
    if (file.exists(filepath)) {
      read_csv(filepath, show_col_types = FALSE)
    } else {
      NULL
    }
  }
  
  list(
    emms            = safe_read("emms.csv"),
    letters         = safe_read("posthoc_letters.csv"),
    model_summary   = safe_read("model_summary.csv"),
    model_contrasts = safe_read("posthoc_contrasts.csv"),
    # Prefer marginal predictions file if available (produced by pipeline),
    # fall back to legacy 'predictions.csv'.
    predictions     = (function() {
      pm <- safe_read("predictions_marginal.csv")
      if (!is.null(pm)) return(pm)
      safe_read("predictions.csv")
    })()
  )
}

# ============================================================
#                 FIGURE 2 DATA
# ============================================================

# Pika
pika_data            <- load_csv("data/clean/diet_selection/pika_feeding.csv")
pika_feeding_models  <- load_model_data("data/clean/diet_selection/pika_feeding")

# Yak
yak_data             <- load_csv("data/clean/diet_selection/yak_grazing.csv")
yak_feeding_models   <- load_model_data("data/clean/diet_selection/yak_grazing")

# Shared dung/burrow
dung_burrow_by_plot  <- load_csv("data/clean/diet_selection/dung_burrow_by_plot.csv") |>
  mutate(plot = factor(plot))

# ============================================================
#                 FIGURE 3 DATA
# ============================================================

# Weight gain
weight_gain <- load_csv("data/clean/foraging_efficiency/yak_weight_gain.csv") |>
  mutate(yak = gsub("_[^_]*$", "", yak))
weight_gain_models <- load_model_data("data/clean/foraging_efficiency/weight_gain")

# Forage quality (all metrics in one pass)
quality_metrics <- c("cp", "adf", "ndf")
forage_quality  <- load_csv("data/clean/forage_quality/plant_forage_quality.csv")
forage_quality_models <- map(quality_metrics,
                             ~ load_model_data("data/clean/forage_quality", .x)) |>
  set_names(quality_metrics)

# Plant cover
cover_data     <- load_csv("data/clean/plant_cover/plant_cover_by_treatment.csv")
grasses_models <- load_model_data("data/clean/plant_cover/grasses")
poison_models  <- load_model_data("data/clean/plant_cover/s_chamaejasme")
forbs_models   <- load_model_data("data/clean/plant_cover/forbs")
sedges_models  <- load_model_data("data/clean/plant_cover/sedges")

# Flip grouping for sedges posthoc letters
sedges_models$letters <- sedges_models$letters |>
  mutate(.group = ifelse(.group == "a", "b", "a"))

# Active burrow count by treatment
act_burrow_count_dat <- read_csv(here('data/clean/additional_data/active_burrow_dat.csv'),show_col_types=FALSE)
burrow_count_models  <- load_model_data("data/clean/additional_data/active_burrow_count")
weight_gain_burrow_regression_models  <- load_model_data("data/clean/additional_data/weight_gain")

# ============================================================
#                 FIGURE 4 DATA (LOADED EARLY FOR INLINE CALCS)
# ============================================================

# Plant bites
plant_bites <- load_csv("data/clean/foraging_efficiency/yak_plant_bites.csv") |>
  mutate(plant = gsub("_[^_]*$", "", plant))
plant_groups <- c("forbs", "grasses", "sedges")
plant_bites_models <- map(plant_groups,
                          ~ load_model_data("data/clean/foraging_efficiency/plant_bites",
                                            .x)) |>
  set_names(plant_groups)

# Bite/steps
bite_steps <- load_csv("data/clean/foraging_efficiency/yak_bite_steps_ratio.csv")
bite_step_groups <- c("forbs", "grasses", "sedges")
bite_step_models <- map(bite_step_groups,
                        ~ load_model_data("data/clean/foraging_efficiency/bite_steps_ratio",
                                          .x)) |>
  set_names(bite_step_groups)

# Weight by burrow count
# Data is pre-filtered by pipeline to pika + S. chamaejasme treatment
weight_burrow_raw <- read_csv(here('data/clean/additional_data/weight_gain_raw.csv'), show_col_types = FALSE)
weight_burrow_modeled <- read_csv(here('data/clean/additional_data/weight_gain_modeled_dat.csv'), show_col_types = FALSE)

# ============================================================
#            SUPPLEMENTARY FIGURES/TABLES DATA
# ============================================================

# Plant cover (reload for supplementary analyses)
forbs_models   <- load_model_data("data/clean/plant_cover/forbs")
sedges_models  <- load_model_data("data/clean/plant_cover/sedges")

# Forage quality
ndf_models     <- load_model_data("data/clean/forage_quality/ndf")

# Yak grazing
yak_feeding_model_summary <- yak_feeding_models$model_summary
yak_feeding_contrasts     <- yak_feeding_models$model_contrasts

```

```{r}
#| label: sample-sizes
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

# Calculate sample sizes for inline reporting
n_pika_plots <- pika_data |> distinct(plot) |> nrow()
n_yak_transects <- yak_data |> distinct(transect) |> nrow()
n_burrow_dung_plots <- dung_burrow_by_plot |> distinct(plot) |> nrow()

# Experiment data - plots and treatment combinations
# Use weight_gain data which has one row per plot
n_experiment_plots <- weight_gain |> 
  distinct(pika_treatment, posion_plant_treatment, block) |> 
  nrow()
n_treatments <- 4  # 2x2 factorial design (pika yes/no × poison yes/no)
n_plots_per_treatment <- n_experiment_plots / n_treatments
n_years_experiment <- 2

# Weight-burrow regression uses filtered data (pika + S. chamaejasme only)
n_weight_burrow_observations <- weight_burrow_raw |> nrow()

```

::: {#fig-hypothesis}
```{r fig-hypothesis}
#| echo: false
#| results: asis
#| message: false
#| warning: false
#| fig-width: 11
#| fig-height: 8

library(cowplot)
library(magick)

hypoth_pic <- png::readPNG(here('data/photos/hypothesis_figure.png'))
hypoth_pic <- ggdraw() + draw_image(hypoth_pic) + theme_void(base_size = base_size)

hypoth_pic


```

The hypothesized facilitative effects of pika on yak in the Tibetan Plateau. In the absence of pika, the poisonous Stellera chamaejasme forbs should exert a strong negative effect on yak growth performance by competing with the palatable grasses and sedges, decreasing their food availability and foraging efficiency (Fig. 1a). However, such negative effects of poisonous plants should be mitigated in the presence of pika, whom eliminate S. chamaejasme forbs by their feeding and clipping activities, leading to a facilitative effect on yak (Fig. 1b). The negative effects of poisonous plants on yak were indicated as black dashed lines, the competition between plant groups were indicated by the black solid lines, and the pathway that pika suppress poisonous plants was indicated by red solid line. The sizes of the lines indicate the strengths of species interactions. Credits: Xiaona Zheng (photographs).
:::

# Methods

## Statistical analyses

All data were analyzed using linear models, generalized linear mixed models, or generalized additive models, with the choice of model and statistical family guided by the structure and distribution of the data. Posthoc comparisons were conducted only when the pika × S. chamaejasme interaction term was significant. For the 2021 field surveys, we fit generalized linear mixed models with plot and month as random effects. We then used generalized additive mixed models for S. chamaejasme cover and active pika burrow density, with plot as a random effect, and linear regression models for dung density and S. chamaejasme cover. For the field manipulation experiments in 2022 and 2023, we constructed generalized linear mixed models with the dependent variables (e.g., grass bites per step, sedge bites, weight gain) regressed against the interactive effect of pika and S. chamaejasme treatments, while including block, year, and month as random effects to capture the hierarchical structure of the data. Models assumed gaussian, beta (for proportions), or tweedie (for non-normal data) distributions, selected based on data type and model fit. A significance threshold of P = 0.05 was applied, with TukeyHSD or Sidak posthoc tests used where appropriate. All data management, modeling, and visualization were carried out in R, with dependencies managed using *renv*. The main modeling packages were *glmmTMB* and *mgcv*, with *DHARMa* used for model diagnostics. Data management relied on the *tidyverse* suite of packages. A complete record of package versions is available in the renv.lock file in the repository: <https://github.com/ddlawton/pika_yak_interactions>.

# Results and Discussion

First, we carried out a set of field observational experiments to investigate diet selection by pika and yak, as well as the associations among pika abundance, S. chamaejasme abundance, and yak activity under natural field conditions. Consistent with previous studies^5,11,24–25^, we found that pika and yak have distinct diet preferences: yaks preferred grasses and sedges, while pika preferred the poisonous S. chamaejasme (@fig-diet-selection A,B, @supptbl-pika-yak-feeding-model-summary, @supptbl-pika-yak-feeding-model-contrast). We also found that S. chamaejasme abundance was negatively associated with active pika burrow abundance (@fig-diet-selection C, @supptbl-pika-burrow-chamaejasme-gam) and with yak foraging activity, as indicated by dung density (@fig-diet-selection D, @supptbl-dung-chamaejasme-lm).

Building on these results, we conducted an in-situ manipulative field experiment using fenced enclosures to test the interactive effects of pika and S. chamaejasme on yak weight gain, foraging quantity and quality, and grazing behavior. For weight gain, yaks gained less in poison-plant plots than in non-poison plots, and there was a significant interaction between pika and S. chamaejasme treatments (@supptbl-yak-performance-model-summary). When pika were present, yaks showed higher weight gain (@supptbl-yak-performance-model-contrast). This effect was driven by pika feeding on S. chamaejasme (@fig-diet-selection B, @supptbl-yak-performance-model-summary, @supptbl-yak-performance-model-contrast), which increased grass cover (@fig-diet-selection C, @supptbl-yak-performance-model-summary, @supptbl-yak-performance-model-contrast) compared to plots with S. chamaejasme but no pika.

```{r}
#| label: forage-qaulity
#| echo: false
#| results: hide
#| message: false
#| warning: false

# Function to calculate % increase from "no pika" to "pika"
calc_pct_increase <- function(emms_df, value_col = "emmean") {
  emms_df |>
    filter(posion_plant_treatment == "S. chamaejasme") |>
    summarise(
      pct_increase = round(
        ( .data[[value_col]][pika_treatment == "pika"] -
          .data[[value_col]][pika_treatment == "no pika"]) /
          .data[[value_col]][pika_treatment == "no pika"] * 100
      )
    ) |>
    pull()
}

cp_increase <- calc_pct_increase(forage_quality_models$cp$emms, value_col = "response")
adf_increase <- calc_pct_increase(forage_quality_models$adf$emms, value_col = "response")


```

Food quantity and quality are key drivers of individual performance and population dynamics in large herbivores^30–32^. In the presence of *S. chamaejasme* forbs, pika doubled the cover of yak’s most preferred grasses (@fig-diet-selection C, @supptbl-yak-performance-model-summary, @supptbl-yak-performance-model-contrast) and increased the crude protein content by approximately `r {cp_increase}`% (@fig-yak-performance D, @supptbl-yak-performance-model-summary, @supptbl-yak-performance-model-contrast), suggesting that pika facilitate yak by enhancing both the quantity and quality of forage. Acid detergent fibre (`r {adf_increase}`%) was higher in pika plots. These increases in grass abundance and forage quality were likely driven by the decline in poisonous plants induced by pika (@fig-yak-performance B), which reduced interspecific competition for shared resources (e.g., light, soil moisture, nutrients) and allowed grasses to expand^27–28^. In the absence of S. chamaejasme forbs, however, the positive effects of pika on food resources and yak weight gain disappeared (@fig-yak-performance A,C,D). The relationships between active burrow count and treatments, as well as between yak weight gain and burrow density, are presented in @fig-yak-performance (panels G and H) and in @supptbl-pika-burrow-count and @supptbl-weight-gain-burrow-regression, respectively.

```{r}
#| label: bite-increase
#| echo: false
#| warning: false
#| message: false
#| results: "hide"


sedge_bite_increase <- calc_pct_increase(plant_bites_models$sedges$emms, value_col = 'emmean')
grass_bite_steps <- calc_pct_increase(bite_step_models$grasses$emms, value_col = 'response')
sedge_bite_steps <- calc_pct_increase(bite_step_models$sedges$emms, value_col = 'response')
  
```

Pika-yak facilitation was also linked to improved foraging efficiency in yak when grazing alongside pika. Optimal foraging theory predicts that animals minimize energy costs to exploit high-quality food items and maximize intake of digestible nutrients, thereby improving performance^33^. In ruminants, bite rate, step rate, and bites per step are key predictors of foraging efficiency^34^ and, ultimately, animal performance^35^. In the presence of *S. chamaejasme* forbs, yak sedge, and grass bite rates increased by roughly `r {sedge_bite_increase}`% (@fig-yak-foraging-efficiency B), and `r {grass_bite_steps}`% (@fig-yak-foraging-efficiency C), respectively, in the treatment where they coexisted with pika (@supptbl-yak-foraging-efficiency-summary, @supptbl-yak-foraging-efficiency-contrast). Yak stepped more often in the presence of *S. chamaejasme* forbs without pika, which meant they consumed more grasses and sedges per step in plots where pika were present.

Sedge bites per step and grass bites per step increased significantly by `r {sedge_bite_steps}`% (@fig-yak-foraging-efficiency E) and `r {grass_bite_steps}`% (@fig-yak-performance F) in plots with pika. These gains in foraging efficiency can be attributed to the decline in *S. chamaejasme* forbs (@fig-yak-performance B), which improved access to palatable grasses and sedges. In this system, increased food availability (@fig-yak-performance C,D) and enhanced foraging efficiency (@fig-yak-performance) likely work together to shape the facilitative effects of pika on yak performance (@fig-yak-performance A).

::: {#fig-diet-selection}
```{r fig-diet-selection}
#| echo: false
#| results: asis
#| message: false
#| warning: false
#| fig-width: 13
#| fig-height: 12
# -------------------------------------------------------------------
# 1. Helper Plot Function
# -------------------------------------------------------------------

library(cowplot)

# Create a feeding/grazing plot (shared structure for pika & yak)
make_feeding_plot <- function(raw_data, emms, letters, yvar, ylabel, 
                              x_order = NULL, box_color = "white", point_color = "black") {
  data <- raw_data |>
    mutate(plant_group = if (!is.null(x_order)) factor(plant_group, levels = x_order) else plant_group)

  ggplot(data, aes(x = plant_group, y = .data[[yvar]])) +
    geom_boxplot(aes(fill = plant_group), color = "black", fill = box_color, alpha = 0.7,linewidth = 1.25) +
    #geom_jitter(width = 0.2, height = 0, color = point_color, alpha = 0.7) +
    geom_text(data = letters, aes(label = .group, y = max(data[[yvar]], na.rm = TRUE) * 1.05), size = geom_text_size) +
    theme_pubr(base_size = base_size) +
    labs(x = NULL, y = ylabel) +
    guides(fill = "none")
}

# -------------------------------------------------------------------
# 2. Load Pre-fitted Models from Saved Files
# -------------------------------------------------------------------

# Load GAM models from saved .qs files (created by targets pipeline)
burrow_model <- qs2::qs_read(here('data/models/burrow_cover_gam.qs'))
dung_model <- qs2::qs_read(here('data/models/dung_cover.qs'))

# Load model summaries and predictions from targets outputs
burrow_model_data <- load_model_data("data/clean/diet_selection/burrow_cover_gam")
dung_model_data <- load_model_data("data/clean/diet_selection/dung_cover")

# Extract GAM smooth estimates for burrow model visualization
burrow_intercept <- coef(burrow_model)[1]
gam_ests <- smooth_estimates(burrow_model) |>
  mutate(adj_est = .estimate + burrow_intercept)

# -------------------------------------------------------------------
# 3. Plots
# -------------------------------------------------------------------

# Feeding/clipping plot (pika)
pika_feeding_plot <- make_feeding_plot(
  raw_data = pika_data,
  emms     = pika_feeding_models$emms,
  letters  = pika_feeding_models$letters,
  yvar     = "feeding_clipping_freq",
  ylabel   = "Feeding/clipping frequency (%)",
  x_order  = c("Grasses","Sedges","Forbs","S. chamaejasme")
)

# Grazing plot (yak)

# Reverse .group letters for yak_feeding_models$letters: Grasses = a, Sedges = b, Forbs = c
yak_feeding_models$letters <- yak_feeding_models$letters |> 
  mutate(.group = case_when(
    plant_group == "Grasses" ~ "a",
    plant_group == "Sedges" ~ "b",
    plant_group == "Forbs" ~ "c",
    TRUE ~ .group
  ))

yak_grazing_plot <- make_feeding_plot(
  raw_data = yak_data,
  emms     = (yak_feeding_models$emms |>
    bind_rows(tibble(
      plant_group = "S. chamaejasme",
      emmean      = 0,
      SE          = 0,
      df          = 0,
      lower.CL    = 0,
      upper.CL    = 0
    ))),
  letters  = yak_feeding_models$letters,
  yvar     = "grazing_freq",
  ylabel   = "Grazing frequency (%)",
  x_order  = c("Grasses","Sedges","Forbs","S. chamaejasme")
) +
  scale_x_discrete(drop = FALSE)

# Burrow density vs S. chamaejasme cover
pika_burrow <- ggplot(dung_burrow_by_plot, aes(x = active_pika_burrow_no_100m2, y = s_chamaejasme_cover_percent)) +
  geom_point(color = "darkgrey") +
  geom_ribbon(data = gam_ests,
              aes(x = active_pika_burrow_no_100m2, ymin = adj_est - .se, ymax = adj_est + .se),
              alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = gam_ests,
            aes(x = active_pika_burrow_no_100m2, y = adj_est),
            color = "blue", linewidth = 1.25) +
  theme_pubr(base_size = base_size) +
  labs(x = expression("Active pika burrow density (no/100 m"^2*")"),
       y = expression(italic("S. chamaejasme")~"cover (%)"))

# Dung density vs cover (using model predictions)
dung_plot <- ggplot(dung_burrow_by_plot, aes(x = s_chamaejasme_cover_percent, y = yak_dung_no_100m2)) +
  geom_point(color = "darkgrey") +
  geom_ribbon(data = dung_model_data$predictions,
              aes(x = s_chamaejasme_cover_percent, ymin = lower, ymax = upper),
              alpha = 0.2, inherit.aes = FALSE) +
  geom_line(data = dung_model_data$predictions,
            aes(x = s_chamaejasme_cover_percent, y = fit),
            color = "blue", linewidth = 1.25) +
  theme_pubr(base_size = base_size) +
  labs(x = expression(italic("S. chamaejasme")~"cover (%)"),
       y = "Yak dung density (no/100 m^2)") +
  coord_cartesian(xlim = c(0, 35), ylim = c(0, 25))

# adding picture of yak
yak_pic <-  png::readPNG(here('data/photos/diet_selection_photo.png'), native = TRUE)
yak_pic <- ggdraw() + draw_image(yak_pic,scale=1) + theme_void(base_size = base_size)



# Combine into Figure 2
layout <- "
AB
AB
DE
DE
"


figure_2_plot <- pika_feeding_plot + 
  yak_grazing_plot +
  pika_burrow + 
  dung_plot + 
  plot_annotation(tag_levels = "A") +
  plot_layout(design = layout, guides = "collect") &
  theme(legend.position = "bottom")

# -------------------------------------------------------------------
# 4. plot Output
# -------------------------------------------------------------------

figure_2_plot

```

Diet selection of pikas and yaks and their potential interactions
mediated by poisonous plants based on field surveys in July 2021. (A) Feeding
and clipping frequencies of pikas (n = `r n_pika_plots` plots), and (B) grazing frequencies of yaks on grasses,
sedges, forbs, and Stellera chamaejasme (n = `r n_yak_transects` transects) across `r n_pika_plots` 2 × 2 m plots and `r n_yak_transects` 250 m
transects, respectively. (C) Relationship between active pika burrow density and S.
chamaejasme cover (n = `r n_burrow_dung_plots` plots), and (D) between S. chamaejasme cover and an index of yak
grazing activity (dung density) in `r n_burrow_dung_plots` 10 × 10 m plots. (E) Photographs showing (left
to right) a pika at its burrow entrance, flowering S. chamaejasme, and a yak grazing
among S. chamaejasme plants. Different letters above bars indicate significant
differences at P &lt; 0.05. Credits: Xiaona Zheng (photographs).
:::

::: {#fig-yak-performance}
```{r}
#| echo: false
#| results: asis
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 12

# Helper to check if interaction is significant
interaction_significant <- function(model_summary, interaction_pattern = "pika.*:.*chamaejasme", alpha = 0.05) {
  # Find the interaction row (term contains both pika and chamaejasme)
  row <- model_summary[grepl(interaction_pattern, model_summary$term), ]
  if (nrow(row) == 0) return(FALSE)
  # Use p.value or p.value column (could be named differently)
  pval_col <- intersect(c("p.value", "pvalue", "p_val", "p.val"), names(row))
  if (length(pval_col) == 0) return(FALSE)
  return(any(row[[pval_col[1]]] < alpha, na.rm = TRUE))
}

# Generic plot for treatment × response with jitter, means, and letters
make_treatment_plot <- function(data, emms, letters, response_var, y_label, label_y = NULL, model_summary = NULL) {
  if ("response" %in% names(emms)) {
    emms <- emms |> mutate(.mean_val = response * 100)
  } else if ("emmean" %in% names(emms)) {
    emms <- emms |> mutate(.mean_val = emmean)
  } else {
    stop("emms must contain either 'response' or 'emmean' column")
  }
  label_y <- label_y %||% max(data[[response_var]], na.rm = TRUE)
  # Only show letters if interaction is significant
  show_letters <- TRUE
  if (!is.null(model_summary)) {
    show_letters <- interaction_significant(model_summary)
  }
  p <- ggplot(data, aes(x = pika_treatment, y = .data[[response_var]], fill = posion_plant_treatment)) +
    geom_boxplot(position = position_dodge(width = 1), color = "black", outlier.shape = NA) +
    scale_fill_manual(values = plant_colors) +
    scale_color_manual(values = plant_colors) +
    labs(x = NULL, y = y_label, color = NULL, fill = NULL) +
    theme_pubr(base_size = base_size) +
    theme(legend.position = "bottom")
  if (!is.null(letters) && show_letters) {
    p <- p +
      geom_text(
        data = letters,
        aes(x = pika_treatment, y = label_y, label = .group, group = posion_plant_treatment),
        position = position_dodge(width = 1),
        size = geom_text_size, color = "black", show.legend = FALSE
      )
  }
  p
}


# Update letter groupings
weight_gain_models$letters <- weight_gain_models$letters |> mutate(.group = ifelse(.group == 'a','b','a'))
grasses_models$letters <- grasses_models$letters |> mutate(.group = ifelse(.group == 'a','b','a'))
forage_quality_models$cp$letters <- forage_quality_models$cp$letters |> mutate(.group = ifelse(.group == 'a','b','a'))
forage_quality_models$adf$letters <- forage_quality_models$adf$letters |> mutate(.group = ifelse(.group == 'a','b','a'))

# Panels (letters only if interaction significant)
panel_a <- make_treatment_plot(
  data    = weight_gain,
  emms    = weight_gain_models$emms,
  letters = weight_gain_models$letters,
  response_var = "weight_gain",
  y_label = "Weight gain (kg/yak/day)",
  label_y = 0.60,
  model_summary = weight_gain_models$model_summary
)

panel_b <- make_treatment_plot(
  data    = cover_data |> filter(plant == "s_chamaejasme"),
  emms    = poison_models$emms,
  letters = poison_models$letters,
  response_var = "cover",
  y_label = expression(italic("S. chamaejasme") ~ "cover (%)"),
  label_y = 43,
  model_summary = poison_models$model_summary
)

panel_c <- make_treatment_plot(
  data    = cover_data |> filter(plant == "grasses"),
  emms    = grasses_models$emms,
  letters = grasses_models$letters,
  response_var = "cover",
  y_label = "Grass cover (%)",
  label_y = 41,
  model_summary = grasses_models$model_summary
)

panel_d <- make_treatment_plot(
  data    = forage_quality |> filter(plant == "cp"),
  emms    = forage_quality_models$cp$emms,
  letters = forage_quality_models$cp$letters,
  response_var = "forage_quality",
  y_label = "CP (%)",
  label_y = 10,
  model_summary = forage_quality_models$cp$model_summary
)

panel_e <- make_treatment_plot(
  data    = forage_quality |> filter(plant == "adf"),
  emms    = forage_quality_models$adf$emms,
  letters = forage_quality_models$adf$letters,
  response_var = "forage_quality",
  y_label = "ADF (%)",
  label_y = 36,
  model_summary = forage_quality_models$adf$model_summary
)

sedge_plot <- make_treatment_plot(
  data    = cover_data |> filter(plant == "sedges"),
  emms    = sedges_models$emms,
  letters = sedges_models$letters,
  response_var = "cover",
  y_label = "Sedge cover (%)",
  label_y = 70,
  model_summary = sedges_models$model_summary
)

ndf_plot <- make_treatment_plot(
  data    = forage_quality |> filter(plant == "ndf"),
  emms    = forage_quality_models$ndf$emms,
  letters = NULL,#forage_quality_models$ndf$letters,
  response_var = "forage_quality",
  y_label = "NDF (%)",
  label_y = 70
)

panel_g <- ggplot(act_burrow_count_dat, aes(x = pika_treatment, y = active_burrows, color = posion_plant_treatment)) +
  geom_boxplot(aes(fill = posion_plant_treatment), position = position_dodge(width = 1), color = "black", outlier.shape = NA) +
  scale_fill_manual(values = plant_colors) +
  scale_color_manual(values = plant_colors) +
  labs(x = NULL, y = 'No. active burrows/ha', color = NULL, fill = NULL) +
  theme_pubr(base_size = base_size) +
  theme(legend.position = "bottom")

panel_h <- weight_burrow_modeled |>
  ggplot(aes(x = active_burrows, y = weight_gain_mean)) +
  geom_point(data = weight_burrow_raw, aes(y = weight_gain_mean), color = plant_colors[2]) +
  geom_line(aes(y = weight_gain_mean), size = 1) +
  theme_pubr() +
  xlab('No. active burrows/ha') +
  ylab('Weight gain kg/yak/day') +
  xlim(0, 500) +
  ylim(0, 0.5)

# 3x3 layout
layout <- "
ABC
DEF
GHI
"

figure_3 <- panel_g + panel_a + panel_h +
            panel_b + panel_c + sedge_plot +
            panel_d + panel_e + plot_spacer() +
            plot_annotation(tag_levels = "A") +
            plot_layout(design = layout, guides = "collect") & theme(legend.position = "bottom")

figure_3

```

Combined effects of two-year (2022–2023) pika and Stellera
chamaejasme removal treatments on yak growth, forage quantity, and quality in
the field manipulative experiment. (A) Pika density (indicated by active burrow
density), (B) yak weight gain, (C) relationship between pika density and yak weight
gain, (D) S. chamaejasme cover, (E) grass cover, (F) sedge cover, (G) crude protein
(CP) content, and (H) acid detergent fibre (ADF) content. Average values from both years were used in
analyses, yielding one data point per 150 × 150 m plot (n = `r n_experiment_plots` plots: `r n_plots_per_treatment` plots per treatment combination). Significant interactions
between pikas and S. chamaejasme were evaluated using post hoc tests; means not
sharing letters differ significantly. For panels A and F, only main effects were
significant (Table S3). Error bars indicate +/- SE.
:::

::: {#fig-yak-foraging-efficiency}
```{r fig-yak-foraging-efficiency}
#| echo: false
#| results: asis
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 8


# ============================================================
#                 PLANT BITE DATA
# ============================================================


# Adjust letter groupings consistently
plant_bites_models <- map(plant_bites_models, function(x) {
  x$letters <- x$letters |> mutate(.group = ifelse(.group == "a", "b", "a"))
  x
})

# ============================================================
#                 BITE/STEPS DATA
# ============================================================

# Adjust letter groupings consistently
bite_step_models <- map(bite_step_models, function(x) {
  x$letters <- x$letters |>
    mutate(.group = ifelse(.group %in% c("a", "A"), "b", "a"))
  x
})

# ============================================================
#                 REUSABLE PLOTTING FUNCTIONS
# ============================================================
# --- Plant Bites Plot (panels A–C) ---
plot_plant_bites <- function(plant_name, y_label, label_y = NULL, letters = TRUE) {
  data <- plant_bites |> filter(plant == plant_name)
  model <- plant_bites_models[[plant_name]]
  # Standardize emms table
  emms <- model$emms
  if ("response" %in% names(emms)) {
    emms <- emms |> mutate(.mean_val = response * 100)
  } else if ("emmean" %in% names(emms)) {
    emms <- emms |> mutate(.mean_val = emmean)
  } else {
    stop("model$emms must contain either 'response' or 'emmean'")
  }
  # Standardize letters table
  letters_df <- model$letters
  if ("response" %in% names(letters_df)) {
    letters_df <- letters_df |> mutate(.mean_val = response)
  } else if ("emmean" %in% names(letters_df)) {
    letters_df <- letters_df |> mutate(.mean_val = emmean)
  } else {
    stop("model$letters must contain either 'response' or 'emmean'")
  }
  # Set default y-position for labels if not supplied
  label_y <- label_y %||% (max(data$forage_efficiency, na.rm = TRUE) + 0.05)
  # Only show letters if interaction is significant
  show_letters <- TRUE
  if (!is.null(model$model_summary)) {
    show_letters <- interaction_significant(model$model_summary)
  }
  p <- ggplot(data, aes(x = pika_treatment, y = forage_efficiency, fill = posion_plant_treatment)) +
    geom_boxplot(position = position_dodge(width = 1), color = "black", outlier.shape = NA) +
    scale_fill_manual(values = plant_colors) +
    labs(
      y = y_label, x = "",
      fill = "Poison plant treatment"
    ) +
    theme_pubr(base_size = base_size) +
    theme(legend.position = "bottom", legend.title = element_blank())
  if (letters && show_letters) {
    p <- p +
      geom_text(
        data = letters_df,
        aes(x = pika_treatment, y = label_y, label = .group, group = posion_plant_treatment),
        position = position_dodge(width = 1),
        size = geom_text_size, color = "black", show.legend = FALSE, inherit.aes = FALSE
      )
  }
  p
}

# --- Bite/Step Ratio Plot (panels E and F) ---
plot_bite_steps <- function(plant_name, y_label, label_y = NULL, add_letters = TRUE) {
  data <- bite_steps |> filter(plant == plant_name)
  model <- bite_step_models[[plant_name]]
  label_y <- label_y %||% max(data$bites_steps_ratio, na.rm = TRUE) + 0.05
  # Only show letters if interaction is significant
  show_letters <- TRUE
  if (!is.null(model$model_summary)) {
    show_letters <- interaction_significant(model$model_summary)
  }
  p <- ggplot(data, aes(x = pika_treatment, y = bites_steps_ratio, fill = posion_plant_treatment)) +
    geom_boxplot(position = position_dodge(width = 1), color = "black", outlier.shape = NA) +
    scale_fill_manual(values = plant_colors) +
    labs(y = y_label, x = "", fill = "Poison plant treatment") +
    theme_pubr(base_size = base_size) +
    theme(legend.position = "bottom", legend.title = element_blank())
  if (add_letters && show_letters) {
    p <- p + geom_text(data = model$letters,
                       aes(x = pika_treatment, y = label_y, label = .group, group = posion_plant_treatment),
                       position = position_dodge(width = 1),
                       size = geom_text_size, color = "black", show.legend = FALSE)
  }
  p
}


# ============================================================
#                 COMBINE PANELS
# ============================================================

# Generate bite panels
panel_a <- plot_plant_bites("sedges", "Sedge bites", label_y = 450)
panel_b <- plot_plant_bites("grasses", "Grass bites", label_y = 350)

# Generate bite/step panels
panel_c <- plot_bite_steps("sedges", "Sedge bite/step", label_y = 1)
panel_d <- plot_bite_steps("grasses", "Grass bite/step")

# Layout specification
layout <- "
AB
CD
"

figure_4 <- panel_a + panel_b + panel_c +
            panel_d +
            plot_annotation(tag_levels = "A") +
            plot_layout(design = layout, guides = "collect") &
            theme(legend.position = "bottom")

figure_4

```

Combined effects of two-year (2022–2023) pika and Stellera
chamaejasme removal treatments on yak foraging efficiency in the field
manipulative experiment. (A, B) Bite rates and (C, D) bites per step for sedges and
grasses, respectively. Mean values from both years were used for analysis (one data
point per 150 × 150 m plot; n = `r n_experiment_plots` plots: `r n_plots_per_treatment` plots per treatment combination). Significant interactions between pikas and S.
chamaejasme were evaluated using post hoc tests; means not sharing letters differ
significantly. Error bars indicate +/- SE.
:::

# Supplementary Figures and Tables

::: {#suppfig-ndf-plot}

```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false
# ndf_plot <- make_treatment_plot(
#   data    = forage_quality |> filter(plant == "ndf"),
#   emms    = forage_quality_models$ndf$emms,
#   letters = NULL,#forage_quality_models$ndf$letters,
#   response_var = "forage_quality",
#   y_label = "NDF (%)",
#   label_y = 70
# )

# ndf_plot

forbs_plot <- make_treatment_plot(
  data    = cover_data |> filter(plant == "forbs"),
  emms    = forbs_models$emms,
  letters = forbs_models$letters,
  response_var = "cover",
  y_label = "Forb cover (%)",
  label_y = 45,
  model_summary = forbs_models$model_summary
)

```

Combined effects of two-year (2022–2023) pika and Stellera
chamaejasme removal on forb cover in the field manipulative experiment. The
average values of each variable in the two years were used for statistical analysis,
providing a single data point for each variable in each 150 × 150 m plot (n = `r n_experiment_plots` plots: `r n_plots_per_treatment` plots per treatment combination). Error bars
represent +/- SE.
:::


::: {#suppfig-forb-bite-steps}

```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

# Generate bite panels
panel_a <- plot_plant_bites("forbs", "Forb bites", label_y = 200,letters=FALSE)

# Generate bite/step panels
panel_b <- plot_bite_steps("forbs", "Forbs bite/steps", label_y = 0.4, add_letters = FALSE)

# Layout specification
layout <- "
AB
"

supp_fig_2 <- panel_a + panel_b + 
            plot_annotation(tag_levels = "A") +
            plot_layout(design = layout, guides = "collect") &
            theme(legend.position = "bottom")

supp_fig_2


```

Combined effects of two-year (2022–2023) pika and Stellera
chamaejasme removal on foraging efficiency of yaks in the field manipulative experiment. (A) bite rates and (B) bites per step of yaks for forbs, respectively. The
average values of each variable in the two years were used for statistical analysis,
providing a single data point for each variable in each 150 × 150 m plot (n = `r n_experiment_plots` plots: `r n_plots_per_treatment` plots per treatment combination). Error bars
represent +/- SE.
:::



::: {#supptbl-pika-yak-feeding-model-summary}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false


bind_rows(
  pika_feeding_models$model_summary |> mutate(model = "Pika feeding"),
  yak_feeding_models$model_summary  |> mutate(model = "Yak feeding")
) |>
  mutate(term = case_when(str_detect(term,'sd__') ~ group,
    TRUE ~ term),
    effect = ifelse(effect == 'ran_pars','random','fixed'),
    term = gsub('plant_group','',term)) |>
  select(-c(component,group)) |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Model summary for pika and yak feeding. Pika feeding model: beta distribution, n = `r n_pika_plots` plots. Yak feeding model: beta distribution, n = `r n_yak_transects` transects.
:::

::: {#supptbl-pika-yak-feeding-model-contrast}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

bind_rows(
  pika_feeding_models$model_contrasts |> mutate(model = "Pika feeding"),
  yak_feeding_models$model_contrasts  |> mutate(model = "Yak feeding")
) |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Model contrasts for pika and yak feeding (pika: n = `r n_pika_plots` plots; yak: n = `r n_yak_transects` transects).
:::

::: {#supptbl-pika-burrow-count}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false


burrow_count_models$model_summary  |>
  mutate(term = case_when(str_detect(term,'sd__') ~ group,
    TRUE ~ term),
    effect = ifelse(effect == 'ran_pars','random','fixed'),
    term = gsub('plant_group','',term)) |>
  dplyr::select(-c(component,group)) |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)

```

Model summary for active burrow count between treatment levels. Estimates are from a generalized linear mixed model (Tweedie distribution) with block, year, and month as random effects (n = `r n_experiment_plots` plots, `r n_years_experiment` years).
:::

::: {#supptbl-pika-burrow-chamaejasme-gam}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

library(broom)

burrow_model_summary <- broom::tidy(burrow_model, parametric = TRUE) |>
  bind_rows(broom::tidy(burrow_model)) |>
  mutate(term = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "s(active_pika_burrow_no_100m2)" ~ "s(active pika burrow density)",
    term == "s(plot)" ~ "s(plot) [random effect]",
    TRUE ~ term
  ))

burrow_model_summary |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 2)

```

Model summary for S. chamaejasme cover regressed on active pika burrow density. Estimates are from a generalized additive mixed model (Gaussian family) with plot as a random effect (n = `r n_burrow_dung_plots` plots).
:::

::: {#supptbl-dung-chamaejasme-lm}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

dung_model_summary <- broom::tidy(dung_model) |>
  mutate(term = case_when(
    term == "(Intercept)" ~ "Intercept",
    term == "s_chamaejasme_cover_percent" ~ "S. chamaejasme cover (%)",
    TRUE ~ term
  ))

dung_model_summary |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 2)

```

Model summary for yak dung density regressed on S. chamaejasme cover. Estimates are from a linear regression model (n = `r n_burrow_dung_plots` plots).
:::

::: {#supptbl-weight-gain-burrow-regression}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false


weight_gain_burrow_regression_models$model_summary  |>
  gt() |>
  fmt_number(columns = where(is.numeric), decimals = 2)

```

Model summary for yak weight gain regressed on active burrow count in plots with both pika presence and S. chamaejasme presence. Estimates are from a generalized additive mixed model (Gaussian family) with block, year, and month as random effects (n = `r n_weight_burrow_observations` observations from the pika + S. chamaejasme treatment).
:::

::: {#supptbl-yak-performance-model-summary}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

yak_performance_summary_table <- bind_rows(
  weight_gain_models$model_summary |> mutate(model = 'weight gain'),
  grasses_models$model_summary |> mutate(model = 'grass cover'),
  sedges_models$model_summary |> mutate(model = 'sedge cover'),
  forbs_models$model_summary |> mutate(model = 'forb cover'),
  poison_models$model_summary |> mutate(model = 'S. chamaejasme'),
  forage_quality_models$cp$model_summary |> mutate(model = 'crude protein %'),
  forage_quality_models$adf$model_summary |> mutate(model = 'acid detergent fibre %'),
  forage_quality_models$ndf$model_summary |> mutate(model = 'neutral detergent fiber %'))  |>
  mutate(term = case_when(str_detect(term,'sd__') ~ group,
    TRUE ~ term),
    effect = ifelse(effect == 'ran_pars','random','fixed'),
    term = gsub('pika_treatment|posion_plant_treatment','',term)) |>
  select(-c(component,group))

yak_performance_summary_table |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)

```

Model summary for weight gain, grass cover, S. chamaejasme cover, crude protein %, acid detergent fibre %. All models are generalized linear mixed models with block, year, and month as random effects (n = `r n_experiment_plots` plots, `r n_years_experiment` years). Distributions: weight gain (Gaussian), cover variables (beta), forage quality variables (Gaussian).
:::

::: {#supptbl-yak-performance-model-contrast}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false


yak_performance_contrast_table <- bind_rows(
  weight_gain_models$model_contrasts |> mutate(model = 'weight gain'),
  grasses_models$model_contrasts |> mutate(model = 'grass cover'),
  poison_models$model_contrasts |> mutate(model = 'S. chamaejasme'),
  forage_quality_models$cp$model_contrasts |> mutate(model = 'crude protein %'),
  forage_quality_models$adf$model_contrasts |> mutate(model = 'acid detergent fibre %'))  |>
  select(contrast,estimate,SE,df,t.ratio,p.value,model)

yak_performance_contrast_table |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Model contrasts for weight gain, grass cover, S. chamaejasme cover, crude protein %, and acid detergent fibre % of total forage for yaks (n = `r n_experiment_plots` plots, `r n_years_experiment` years).

:::

::: {#supptbl-yak-foraging-efficiency-summary}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

yak_foraging_efficiency_table <- bind_rows(
  plant_bites_models$sedges$model_summary |> mutate(model = 'sedge bites/h'),
  plant_bites_models$grasses$model_summary |> mutate(model = 'grass bites/h'),
  plant_bites_models$forbs$model_summary |> mutate(model = 'forb bites/h'),
  bite_step_models$sedges$model_summary |> mutate(model = 'sedges bite steps'),
  bite_step_models$grasses$model_summary |> mutate(model = 'grass bite steps'),
  bite_step_models$forbs$model_summary |> mutate(model = 'forb bite steps')
  )    |>
  mutate(term = case_when(str_detect(term,'sd__') ~ group,
    TRUE ~ term),
    effect = ifelse(effect == 'ran_pars','random','fixed'),
    term = gsub('pika_treatment|posion_plant_treatment','',term)) |>
  select(-c(component,group))

yak_foraging_efficiency_table |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Model summary for plant bites/h and the bites to step ratio for sedges, grasses, and forbs. All models are generalized linear mixed models (beta distribution for proportions, Gaussian for counts) with block, year, and month as random effects (n = `r n_experiment_plots` plots, `r n_years_experiment` years).
:::

::: {#supptbl-yak-foraging-efficiency-contrast}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

yak_foraging_efficiency_table <- bind_rows(
  plant_bites_models$sedges$model_contrasts |> mutate(model = 'sedge bites/h'),
  plant_bites_models$grasses$model_contrasts |> mutate(model = 'grass bites/h'),
  bite_step_models$sedges$model_contrasts |> mutate(model = 'sedges bite steps'),
  bite_step_models$grasses$model_contrasts |> mutate(model = 'grass bite steps')
  )  |>
  select(contrast,estimate,SE,df,t.ratio,p.value,model)

yak_foraging_efficiency_table |>
  gt(groupname_col = "model") |>
  fmt_number(columns = where(is.numeric), decimals = 2)


```

Model contrasts for plant bites and the bites to step ratio for sedges, grasses, and forbs (n = `r n_experiment_plots` plots, `r n_years_experiment` years).
:::

::: {#supptbl-model-overview}
```{r}
#| echo: false
#| results: 'asis'
#| message: false
#| warning: false

# ============================================================
#          HELPER FUNCTIONS TO EXTRACT MODEL INFO
# ============================================================

# Extract response variable from model formula
get_response_var <- function(model) {
  if (is.null(model)) return(NA_character_)
  
  # Handle adaptive models (fit_plant_bites_model returns a list)
  if (is.list(model) && "model" %in% names(model)) {
    model <- model$model
  }
  
  if (is.null(model)) return(NA_character_)
  
  formula_obj <- tryCatch(formula(model), error = function(e) NULL)
  if (is.null(formula_obj)) return(NA_character_)
  
  # Get left-hand side of formula
  response <- tryCatch({
    as.character(formula_obj[[2]])
  }, error = function(e) {
    # For some model types, try alternative extraction
    if (inherits(model, "gam")) {
      # Get response from model terms
      terms_obj <- terms(model)
      response_idx <- attr(terms_obj, "response")
      if (response_idx > 0) {
        all_vars <- all.vars(terms_obj)
        return(all_vars[response_idx])
      }
    }
    return(NA_character_)
  })
  
  return(response)
}

# Extract fixed effects (interaction terms)
get_fixed_effects <- function(model) {
  if (is.null(model)) return(NA_character_)
  
  # Handle adaptive models
  if (is.list(model) && "model" %in% names(model)) {
    model <- model$model
  }
  
  if (is.null(model)) return(NA_character_)
  
  # For GAM models
  if (inherits(model, "gam")) {
    formula_obj <- formula(model)
    # Extract terms, excluding smooths and response
    terms_char <- as.character(formula_obj)[3]
    # Look for interaction patterns
    if (grepl(":", terms_char) || grepl("\\*", terms_char)) {
      # Extract interaction terms
      fixed_terms <- gsub("s\\([^)]+\\)", "", terms_char)
      fixed_terms <- gsub("\\+\\s*\\+", "+", fixed_terms)
      fixed_terms <- trimws(fixed_terms)
      return(fixed_terms)
    } else {
      return("")
    }
  }
  
  # For glmmTMB models
  if (inherits(model, "glmmTMB")) {
    # Get model summary
    summ <- summary(model)
    coef_table <- summ$coefficients$cond
    
    # Look for interaction terms (contain ":")
    if (!is.null(coef_table)) {
      row_names <- rownames(coef_table)
      interaction_terms <- row_names[grepl(":", row_names)]
      
      if (length(interaction_terms) > 0) {
        # Clean up the interaction term names - extract just the base variable names
        interactions <- sapply(interaction_terms, function(term) {
          parts <- strsplit(term, ":")[[1]]
          # Clean each part - extract just the variable name, not the factor level
          parts <- gsub("pika_treatment.*", "pika", parts)
          parts <- gsub("posion_plant_treatment.*", "poison", parts)
          paste(parts, collapse = " × ")
        })
        return(paste(unique(interactions), collapse = "; "))
      }
    }
  }
  
  # For lm models
  if (inherits(model, "lm")) {
    # Check if there are interaction terms
    terms_obj <- terms(model)
    term_labels <- attr(terms_obj, "term.labels")
    interaction_terms <- term_labels[grepl(":", term_labels)]
    
    if (length(interaction_terms) > 0) {
      return(paste(interaction_terms, collapse = "; "))
    }
  }
  
  return("")
}

# Extract random effects
get_random_effects <- function(model) {
  if (is.null(model)) return(NA_character_)
  
  # Handle adaptive models
  if (is.list(model) && "model" %in% names(model)) {
    model <- model$model
  }
  
  if (is.null(model)) return(NA_character_)
  
  # For GAM models
  if (inherits(model, "gam")) {
    formula_obj <- formula(model)
    terms_char <- as.character(formula_obj)[3]
    # Extract random effect smooths (bs = 're')
    re_matches <- gregexpr("s\\(([^)]+),\\s*bs\\s*=\\s*['\"]re['\"]\\)", terms_char, perl = TRUE)
    if (re_matches[[1]][1] != -1) {
      re_text <- regmatches(terms_char, re_matches)[[1]]
      # Extract variable names
      re_vars <- gsub("s\\(([^,]+),.+", "\\1", re_text)
      return(paste(re_vars, collapse = ", "))
    }
    # Check if there are any random effects in the model
    smooth_terms <- summary(model)$s.table
    if (!is.null(smooth_terms) && nrow(smooth_terms) > 0) {
      row_names <- rownames(smooth_terms)
      # Look for "re" basis functions
      re_terms <- row_names[grepl("s\\(.+\\)", row_names) & grepl(",re\\)", row_names)]
      if (length(re_terms) > 0) {
        # Extract variable names from smooth terms like "s(plot)"
        re_vars <- gsub("s\\(([^,)]+).*\\)", "\\1", re_terms)
        return(paste(re_vars, collapse = ", "))
      }
    }
    return("")
  }
  
  # For glmmTMB models
  if (inherits(model, "glmmTMB")) {
    re_struct <- model$modelInfo$reTrms
    if (!is.null(re_struct) && length(re_struct$cond$cnms) > 0) {
      re_names <- names(re_struct$cond$cnms)
      return(paste(re_names, collapse = ", "))
    }
  }
  
  # For lm models (no random effects)
  if (inherits(model, "lm") && !inherits(model, "glmmTMB")) {
    return("")
  }
  
  return("")
}

# Extract family and Tweedie power
get_family_info <- function(model) {
  if (is.null(model)) return(list(family = NA_character_, tweedie_power = NA_real_))
  
  # Handle adaptive models - the actual fitted model is in $model
  # But only if this is truly an adaptive model wrapper (not a GAM/LM with a $model frame)
  actual_model <- model
  if (is.list(model) && "model" %in% names(model) && 
      !inherits(model, c("lm", "glm", "gam", "glmmTMB"))) {
    actual_model <- model$model
  }
  
  if (is.null(actual_model)) return(list(family = NA_character_, tweedie_power = NA_real_))
  
  # Get family
  fam <- tryCatch({
    if (inherits(actual_model, "glmmTMB")) {
      # Extract family from glmmTMB model object directly
      fam_obj <- actual_model$modelInfo$family
      if (!is.null(fam_obj)) {
        list(
          family = fam_obj$family,
          link = fam_obj$link
        )
      } else {
        NULL
      }
    } else {
      family(actual_model)
    }
  }, error = function(e) NULL)
  
  if (is.null(fam)) {
    # For lm objects
    if (inherits(actual_model, "lm")) {
      return(list(family = "Gaussian", tweedie_power = NA_real_))
    }
    return(list(family = NA_character_, tweedie_power = NA_real_))
  }
  
  family_name <- fam$family
  if (is.null(family_name) || length(family_name) == 0) {
    family_name <- NA_character_
  }

  # Format family name
  if (!is.na(family_name) && grepl("tweedie", family_name, ignore.case = TRUE)) {
    # Extract Tweedie power parameter
    tweedie_power <- NA_real_
    
    if (inherits(actual_model, "glmmTMB")) {
      # Try multiple ways to extract the power parameter
      tweedie_power <- tryCatch({
        # Method 1: from family object
        if (!is.null(fam$par)) {
          return(fam$par)
        }
        # Method 2: from model info
        if (!is.null(actual_model$modelInfo$family$par)) {
          return(actual_model$modelInfo$family$par)
        }
        # Method 3: from family link parameters
        if (!is.null(actual_model$fit$par)) {
          # The Tweedie power is stored in the dispersion parameters
          disp_par <- actual_model$fit$par
          if (length(disp_par) >= 2) {
            return(disp_par[2])  # Tweedie power is typically the 2nd parameter
          }
        }
        NA_real_
      }, error = function(e) NA_real_)
    }
    
    return(list(family = as.character("Tweedie"), tweedie_power = as.numeric(tweedie_power)))
  } else if (grepl("beta", family_name, ignore.case = TRUE)) {
    return(list(family = as.character("Beta"), tweedie_power = NA_real_))
  } else if (grepl("gaussian", family_name, ignore.case = TRUE)) {
    return(list(family = as.character("Gaussian"), tweedie_power = NA_real_))
  } else if (grepl("binomial", family_name, ignore.case = TRUE)) {
    return(list(family = as.character("Binomial"), tweedie_power = NA_real_))
  } else {
    return(list(family = as.character(ifelse(is.na(family_name), NA_character_, family_name[1])), tweedie_power = NA_real_))
  }
}

# Get model type
get_model_type <- function(model) {
  if (is.null(model)) return(NA_character_)
  
  # Handle adaptive models
  if (is.list(model) && "model" %in% names(model)) {
    model <- model$model
  }
  
  if (is.null(model)) return(NA_character_)
  
  # Check class hierarchy
  model_classes <- class(model)
  
  if (inherits(model, "gam")) {
    return("GAM")
  } else if (inherits(model, "glmmTMB")) {
    return("GLMM")
  } else if (inherits(model, "lm") && !inherits(model, "glm")) {
    return("LM")
  } else if (inherits(model, "glm")) {
    return("GLM")
  } else {
    # Return first class as fallback
    return(paste(model_classes, collapse = "/"))
  }
}

# ============================================================
#          LOAD ALL MODEL OBJECTS FROM .QS FILES
# ============================================================

model_files <- list(
  # Diet selection models
  pika_feeding = here("data/models/pika_feeding.qs"),
  yak_grazing = here("data/models/yak_grazing.qs"),
  burrow_cover_gam = here("data/models/burrow_cover_gam.qs"),
  dung_cover = here("data/models/dung_cover.qs"),
  
  # Weight gain
  weight_gain = here("data/models/weight_gain.qs"),
  weight_gain_burrow_gam = here("data/models/weight_gain_burrow_gam.qs"),
  
  # Plant cover
  plant_cover_grasses = here("data/models/plant_cover_grasses.qs"),
  plant_cover_sedges = here("data/models/plant_cover_sedges.qs"),
  plant_cover_forbs = here("data/models/plant_cover_forbs.qs"),
  plant_cover_s_chamaejasme = here("data/models/plant_cover_s_chamaejasme.qs"),
  active_burrows = here("data/models/active_burrows.qs"),
  
  # Forage quality
  forage_quality_cp = here("data/models/forage_quality_cp.qs"),
  forage_quality_adf = here("data/models/forage_quality_adf.qs"),
  forage_quality_ndf = here("data/models/forage_quality_ndf.qs"),
  
  # Plant bites
  plant_bites_sedges = here("data/models/plant_bites_sedges.qs"),
  plant_bites_grasses = here("data/models/plant_bites_grasses.qs"),
  plant_bites_forbs = here("data/models/plant_bites_forbs.qs"),
  
  # Bite steps ratio
  bite_steps_sedges = here("data/models/bite_steps_ratio_sedges.qs"),
  bite_steps_grasses = here("data/models/bite_steps_ratio_grasses.qs"),
  bite_steps_forbs = here("data/models/bite_steps_ratio_forbs.qs")
)

# Load models with error handling
models <- map(model_files, ~ {
  if (file.exists(.x)) {
    tryCatch({
      obj <- qs2::qs_read(.x)
      # Verify it's actually a model object or a list containing one
      if (is.null(obj)) return(NULL)
      
      # FIRST: Restore class for GAM/LM models that lost their class attributes
      # This must happen BEFORE checking for adaptive models, because GAM/LM objects
      # have a $model component (the model frame) that can be confused with adaptive model structure
      if (is.list(obj) && "formula" %in% names(obj) && "coefficients" %in% names(obj)) {
        # Check for GAM-specific components
        if ("smooth" %in% names(obj) && "sp" %in% names(obj)) {
          class(obj) <- c("gam", "glm", "lm")
        } 
        # Check for LM-specific components (both qr and terms are reliably present)
        else if ("qr" %in% names(obj) && "terms" %in% names(obj)) {
          class(obj) <- "lm"
        }
        # Check for GLM-specific components
        else if ("family" %in% names(obj) && "qr" %in% names(obj)) {
          class(obj) <- c("glm", "lm")
        }
      }
      
      # Check if it's a data.frame masquerading as a model
      if (is.data.frame(obj) && !inherits(obj, c("lm", "glm", "gam", "glmmTMB"))) {
        warning(paste("Model file", basename(.x), "contains a data.frame, not a model object"))
        return(NULL)
      }
      return(obj)
    }, error = function(e) {
      warning(paste("Error loading model from", basename(.x), ":", e$message))
      return(NULL)
    })
  } else {
    warning(paste("Model file not found:", basename(.x)))
    NULL
  }
})

# ============================================================
#          EXTRACT MODEL INFORMATION
# ============================================================

# Debug: Check which models loaded successfully
models_loaded <- map_lgl(models, ~ !is.null(.x))
if (any(!models_loaded)) {
  warning(paste("Failed to load models:", 
                paste(names(models)[!models_loaded], collapse = ", ")))
}

model_info <- tibble(
  model_name = names(models),
  model_object = models
) |>
  mutate(
    response_var = map_chr(model_object, get_response_var),
    interaction_terms = map_chr(model_object, get_fixed_effects),
    random_effects = map_chr(model_object, get_random_effects),
    family_info = map(model_object, get_family_info),
    family = map_chr(family_info, "family"),
    tweedie_power = map_dbl(family_info, "tweedie_power"),
    model_type = map_chr(model_object, get_model_type)
  ) |>
  select(-family_info, -model_object)

# ============================================================
#          CREATE MODEL OVERVIEW TABLE
# ============================================================

# Define analysis categories and sample sizes
model_overview <- model_info |>
  mutate(
    # Clean up variable names for display
    response_display = case_when(
      response_var == "feeding_clipping_freq" ~ "Pika feeding frequency",
      response_var == "grazing_freq" ~ "Yak grazing frequency",
      response_var == "s_chamaejasme_cover_percent" ~ "S. chamaejasme cover",
      response_var == "yak_dung_no_100m2" ~ "Yak dung density",
      response_var == "weight_gain" ~ "Yak weight gain",
      response_var == "weight_gain_mean" ~ "Yak weight gain (vs burrow density)",
      model_name == "burrow_cover_gam" ~ "S. chamaejasme cover (vs burrow density)",
      model_name == "dung_cover" ~ "Yak dung density (vs S. chamaejasme)",
      model_name == "weight_gain_burrow_gam" ~ "Yak weight gain (vs burrow density)",
      response_var == "cover_beta" ~ str_to_title(gsub("plant_cover_", "", model_name)),
      response_var == "active_burrows" ~ "Active burrow density",
      response_var == "forage_quality_beta" ~ case_when(
        grepl("_cp", model_name) ~ "Crude protein (%)",
        grepl("_adf", model_name) ~ "Acid detergent fiber (%)",
        grepl("_ndf", model_name) ~ "Neutral detergent fiber (%)",
        TRUE ~ "Forage quality"
      ),
      response_var == "forage_efficiency" ~ case_when(
        grepl("sedges", model_name) ~ "Sedge bite rate",
        grepl("grasses", model_name) ~ "Grass bite rate",
        grepl("forbs", model_name) ~ "Forb bite rate",
        TRUE ~ "Plant bite rate"
      ),
      response_var == "bites_steps_ratio_beta" ~ case_when(
        grepl("sedges", model_name) ~ "Sedge bites/step",
        grepl("grasses", model_name) ~ "Grass bites/step",
        grepl("forbs", model_name) ~ "Forb bites/step",
        TRUE ~ "Bites/step ratio"
      ),
      TRUE ~ response_var
    ),
    
    # Analysis category
    analysis_category = case_when(
      model_name %in% c("pika_feeding", "yak_grazing", "burrow_cover_gam", "dung_cover") ~ "Diet selection",
      model_name %in% c("weight_gain", "weight_gain_burrow_gam") ~ "Yak performance",
      grepl("plant_cover", model_name) | model_name == "active_burrows" ~ "Plant cover",
      grepl("forage_quality", model_name) ~ "Forage quality",
      grepl("plant_bites|bite_steps", model_name) ~ "Foraging efficiency",
      TRUE ~ "Other"
    ),
    
    # Override model type for known GAM and LM models if detection failed
    model_type = case_when(
      model_name %in% c("burrow_cover_gam", "weight_gain_burrow_gam") & (is.na(model_type) | model_type == "data.frame") ~ "GAM",
      model_name == "dung_cover" & (is.na(model_type) | model_type == "data.frame") ~ "LM",
      TRUE ~ model_type
    ),
    
    # Override family for known models if detection failed
    family = case_when(
      model_name %in% c("burrow_cover_gam", "weight_gain_burrow_gam", "dung_cover") & (is.na(family) | family == "NA") ~ "Gaussian",
      TRUE ~ family
    ),
    
    # Override random effects for known GAM models if detection failed
    random_effects = case_when(
      model_name == "burrow_cover_gam" & (is.na(random_effects) | random_effects == "") ~ "plot",
      model_name == "weight_gain_burrow_gam" & (is.na(random_effects) | random_effects == "") ~ "block, year, month",
      model_name == "dung_cover" & (is.na(random_effects) | random_effects == "") ~ "",
      TRUE ~ random_effects
    ),
    
    # Sample size
    sample_size = case_when(
      model_name == "pika_feeding" ~ glue("n = {n_pika_plots} plots"),
      model_name == "yak_grazing" ~ glue("n = {n_yak_transects} transects"),
      model_name %in% c("burrow_cover_gam", "dung_cover") ~ glue("n = {n_burrow_dung_plots} plots"),
      model_name == "weight_gain_burrow_gam" ~ glue("n = {n_experiment_plots} plots"),
      TRUE ~ glue("n = {n_experiment_plots} plots, {n_years_experiment} years")
    ),
    
    # Format family display with Tweedie power if applicable
    family_display = if_else(
      !is.na(tweedie_power),
      glue("{family} (p = {round(tweedie_power, 2)})"),
      family
    ),
    
    # Clean up interaction terms display
    interaction_display = if_else(
      interaction_terms == "" | is.na(interaction_terms),
      "",
      # Replace the simplified terms from get_fixed_effects
      gsub("poison", "S. chamaejasme", interaction_terms)
    ),
    
    # Clean up random effects display
    random_effects_display = if_else(
      random_effects == "" | is.na(random_effects),
      "",
      random_effects
    )
  ) |>
  select(
    Analysis = analysis_category,
    `Response Variable` = response_display,
    `Model Type` = model_type,
    `Distribution/Family` = family_display,
    `Interaction Terms` = interaction_display,
    `Random Effects` = random_effects_display,
    `Sample Size` = sample_size
  )

# Display table
model_overview |>
  gt(groupname_col = "Analysis") |>
  tab_header(
    title = "Statistical Models Summary",
    subtitle = "Overview of all models used in the study"
  ) |>
  cols_label(
    `Response Variable` = "Response Variable",
    `Model Type` = "Model Type",
    `Distribution/Family` = "Distribution/Family",
    `Interaction Terms` = "Interaction Terms",
    `Random Effects` = "Random Effects",
    `Sample Size` = "Sample Size"
  ) |>
  tab_footnote(
    footnote = "GLMM = Generalized Linear Mixed Model; GAM = Generalized Additive Mixed Model; LM = Linear Model. For Tweedie distributions, p indicates the power parameter."
  ) |>
  opt_stylize(style = 6, color = "gray") |>
  cols_width(
    `Response Variable` ~ px(200),
    `Model Type` ~ px(80),
    `Distribution/Family` ~ px(150),
    `Interaction Terms` ~ px(180),
    `Random Effects` ~ px(150),
    `Sample Size` ~ px(180)
  )

```

Summary of all statistical models used in the study, including response variables, model type, distribution family (with Tweedie power parameter where applicable), interaction terms, random effects structure, and sample sizes. All field experiments (2022-2023) used a 2×2 factorial design with pika presence/absence and S. chamaejasme presence/absence treatments.
:::