---
title: "Pika-Yak Interctions Results Summary"
author: "Douglas Lawton"
date: today
project:
  type: default
  output-dir: ../
format: 
  pdf:
    toc: true
    toc-depth: 2
    number-sections: true
    keep-tex: true
    fig-cap-location: bottom
    documentclass: article
    geometry: margin=1in
    fontsize: 11pt
    highlight-style: pygments
    code-block-bg: true
    code-line-numbers: false
    citation-package: natbib
    colorlinks: true
    linkcolor: blue
    urlcolor: blue
    fig-width: 8
    fig-height: 8
editor: visual
---

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

library(tidyverse)
library(ggpubr)
library(patchwork)
library(here)
library(mgcv)
library(knitr)
library(kableExtra)

# --- Set Root Directory for Project ---
here::i_am("README.md")  # Anchor project for reproducible paths

# --- Custom Color Palette ---
pink <- '#ff8080'
green <- '#2fc09b'

```



# Introduction

This document summarizes the final statistical models, diagnostics, and effect estimates used in the manuscript titled:  
**_"Pika facilitates yaks by suppressing poisonous plants in the Tibetan Plateau"_**.

All analyses were conducted in R using `glmmTMB`, `emmeans`, and other tidyverse tools. Figures and contrasts are shown below for all major response variables.

All code and data are housed within the project’s GitHub repository, available at: [https://github.com/ddlawton/pika_yak_interactions](https://github.com/ddlawton/pika_yak_interactions)

---

Instead of grouping the models by figure number, I organized them by data type, as follows:

**Diet Selection**
- Pika feeding
- Yak grazing
- Dung density
- Pika burrow density

**Experiment: Plant Cover**
- Forbs
- Grasses
- *S. chamaejasme*
- Sedges

**Experiment: Foraging Efficiency**
- Yak bite/step ratios
- Yak bites by type
- Yak total steps
- Yak weight gain

**Experiment: Foraging Quality**
- ADF
- CP
- EE
- NDF

Each variable gets its own model, and the modeling approach is described below.

---

# Overview of Statistical Models

The goal of the statistical approach was to respect the hierarchical structure of the data without overfitting. Because the design was fully factorial, model construction was relatively straightforward. The primary variation across models was whether they included a random effect for yak.


**Without individual yak**

```{r}
#| label: model formula 1
#| eval: false
response ~ pika_treatment * poison_plant_treatment +
    (1 | block) + 
    (1 | year) + 
    (1 | month)
```

**With individual yak**

```{r}
#| label: model formula 2
#| eval: false
response ~ pika_treatment * poison_plant_treatment +
    (1 | block) + 
    (1 | year) + 
    (1 | month) + 
    (1 | yak)
```

## Statistical Families Used

| Family   | When Used                                              |
|----------|--------------------------------------------------------|
| Gaussian | Default, used when residuals appeared homoscedastic   |
| Beta     | Used for percent/proportion data (if converged well)  |
| Tweedie  | Used when overdispersion or skew was a concern         |

In general, percent data was modeled using the Beta family when possible. If model fit or convergence was poor, I used the gaussian family instead. As a fallback, the tweedie distribution was helpful in handling data with zero inflation, overdispersion, or non-normal characteristics.

Random effect terms like `block`, `year`, and `month` often had minimal influence on fixed effect estimates. However, I chose to retain them in the models to reflect the experimental design and sampling structure. Even when the variance explained was negligible, their inclusion helped preserve the integrity of the hierarchical layout and avoid pseudoreplication.

---

# Data Modeling Summary

## Diet Selection

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false


pika_data <- read_csv(here('data/processed/diet_selection/pika_feeding.csv'))
pika_feeding_model_summary <- read_csv(here('data/processed/diet_selection/modeled_data/pika_feeding/model_summary.csv'))
pika_feeding_contrasts <- read_csv(here('data/processed/diet_selection/modeled_data/pika_feeding/posthoc_contrasts.csv'))
pika_feeding_emms <- read_csv(here('data/processed/diet_selection/modeled_data/pika_feeding/emms.csv'))
pika_feeding_letters <- read_csv(here('data/processed/diet_selection/modeled_data/pika_feeding/posthoc_letters.csv'))

yak_data <- read_csv(here('data/processed/diet_selection/yak_grazing.csv'))
yak_feeding_model_summary <- read_csv(here('data/processed/diet_selection/modeled_data/yak_grazing/model_summary.csv'))
yak_feeding_contrasts <- read_csv(here('data/processed/diet_selection/modeled_data/yak_grazing/posthoc_contrasts.csv'))
yak_feeding_emms <- read_csv(here('data/processed/diet_selection/modeled_data/yak_grazing/emms.csv'))
yak_feeding_letters <- read_csv(here('data/processed/diet_selection/modeled_data/yak_grazing/posthoc_letters.csv'))

dung_burrow_by_plot <- read_csv(here('data/processed/diet_selection/dung_burrow_by_plot.csv')) |>
    mutate(plot = factor(plot))

burrow_model <- gam(s_chamaejasme_cover_percent ~
  s(active_pika_burrow_no_100m2, k = 8, bs = 'ts') +
    s(plot, bs = 're'),
  data = dung_burrow_by_plot,
  family = gaussian(),
  select = TRUE
)

gam_ests <- gratia::smooth_estimates(burrow_model)
gam_ests$adj_est <- gam_ests$.estimate + coef(burrow_model)[1]

```


The following figure (@fig-diet_selection_figure) shows pika feedind frequency (A) and yak grazig frequebcy (B) by every pika and posion plant treatment with sigificance shown with letters. It also shows the relationship between pika burrow density (C) and yak dung density (D) by *S. chamaejasme* cover.



```{r}
#| label: fig-diet_selection_figure
#| fig-cap: "diet selection"
#| echo: false
#| reulsts: asis'
#| message: false
#| warning: false

pika_feeding_plot <- pika_data |>
  left_join(pika_feeding_emms, by = 'plant_group') |>
  ggplot(aes(x = reorder(plant_group, emmean), y = feeding_clipping_freq)) +
  geom_jitter(width = 0.2, height = 0) +
  geom_point(aes(y = emmean), size = 5, color = 'black', pch = 21, fill = 'dark red') +
  geom_text(data = pika_feeding_letters, aes(label = .group, y = 105)) +
  theme_pubr() +
  ylab('feeding/clipping frequency (%)') +
  xlab('')

yak_grazing_plot <- yak_data |>
  left_join(yak_feeding_emms, by = 'plant_group') |>
  mutate(plant_group = factor(plant_group, levels = c('S. chamaejasme', 'Forbs', 'Sedges', 'Grasses'))) |>
  ggplot(aes(x = plant_group, y = grazing_freq)) +
  geom_jitter(width = 0.2, height = 0) +
  geom_point(aes(y = emmean), size = 5, color = 'black', pch = 21, fill = 'dark red') +
  geom_text(data = yak_feeding_letters, aes(label = .group, y = 105)) +
  theme_pubr() +
  ylab('grazing frequency (%)') +
  xlab('') +
  scale_x_discrete(breaks = c('S. chamaejasme', 'Forbs', 'Sedges', 'Grasses'), drop = FALSE)

pika_burrow <- dung_burrow_by_plot |>
  ggplot(aes(x = active_pika_burrow_no_100m2, y = s_chamaejasme_cover_percent)) +
  geom_point() +
  geom_line(data = gam_ests, aes(y = adj_est), linewidth = 1.25, color = 'blue') +
  theme_pubr() +
  xlab('active pika burrow density (no/100m^2)') +
  ylab('S. chamaejasme cover (%)')

dung_plot <- dung_burrow_by_plot |>
  ggplot(aes(x = s_chamaejasme_cover_percent, y = yak_dung_no_100m2)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE) +
  theme_pubr() +
  ylab('Dung density (no/100m^2)') +
  xlab('S. chamaejasme cover (%)') +
  ylim(0, 25) +
  xlim(0, 35)


(pika_feeding_plot + yak_grazing_plot) / (pika_burrow + dung_plot) + plot_annotation(tag_levels = 'A') + plot_layout(guides = "collect") & theme(legend.position = 'bottom')


```


### Pika feeding results

There was a signficant effect of plant group on pike feeding (@tbl-pika-feeding-model-summary) with *S. chamaejasme* consumption being the highest,followed by forbs, and then sedges and grasses (@tbl-pika-feeding-model-contrasts)


```{r}
#| label: tbl-pika-feeding-model-summary
#| tbl-cap: "Model summary for pika feeding."
#| echo: false
#| reulsts: asis'
#| message: false
#| warning: false

library(knitr)
library(kableExtra)

pika_feeding_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 2) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-pika-feeding-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for yak grazing by plant group."
#| echo: false
#| reulsts: 'asis'
#| message: false
#| warning: false
library(knitr)
library(kableExtra)

pika_feeding_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))


```


### Yak Grazing results

There was a signficant effect of plant group on yak grazing (@tbl-yak-feeding-model-summary) with grass consumption being the highest,followed by sedges, and then forbs (@tbl-yak-feeding-model-contrasts).


```{r}
#| label: tbl-yak-feeding-model-summary
#| tbl-cap: "Model summary for yak grazing by plant group"
#| echo: false
#| reulsts: asis'
#| message: false
#| warning: false

library(knitr)
library(kableExtra)

yak_feeding_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 2) |>
  kable_styling(latex_options = c("striped"))
```



```{r}
#| label: tbl-yak-feeding-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for yak grazing by plant group"
#| echo: false
#| reulsts: 'asis'
#| message: false
#| warning: false
library(knitr)
library(kableExtra)

yak_feeding_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3, 
        caption = "Estimated marginal means contrasts for pika feeding by plant group.") |>
  kable_styling(latex_options = c("striped"))


```

### burrow and dung density

I did not run specific statistical models for these plots (C and D). However I did notice an interesting non-linear relationship betwen pika burrow density and *S. chamajasme* cover. I think these plots are straight forward and more indepth tables were not needed.

## Experiment: Plant cover

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false
cover_data <- read_csv(here('data/processed/experiment_plant_cover/plant_cover_by_treatment.csv'))

forbs_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/emms.csv'))
forbs_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/posthoc_letters.csv'))
forbs_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/model_summary.csv'))
forbs_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/forbs_cover/posthoc_contrasts.csv'))

grasses_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/emms.csv'))
grasses_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/posthoc_letters.csv'))
grasses_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/model_summary.csv'))
grasses_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/grasses_cover/posthoc_contrasts.csv'))

poison_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/emms.csv'))
poison_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/posthoc_letters.csv'))
poison_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/model_summary.csv'))
poison__model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/s_chamaejasme_cover/posthoc_contrasts.csv'))

sedges_emms <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/emms.csv'))
sedges_letters <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/posthoc_letters.csv'))
sedges_model_summary  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/model_summary.csv'))
sedges_model_contrasts  <- read_csv(here('data/processed/experiment_plant_cover/modeled_data/sedges_cover/posthoc_contrasts.csv'))

```


```{r}
#| label: fig-plant-cover
#| fig-cap: "Plant cover between experimental plots"
#| fig-width: 8
#| fig-height: 8
#| echo: false
#| reulsts: asis
#| message: false
#| warning: false

make_cover_plot <- function(plant_name, letters_df, y_label, text_y = 41) {
  cover_data |>
    filter(plant == plant_name) |>
    left_join(letters_df, by = c("pika_treatment", "posion_plant_treatment")) |>
    ggplot(aes(x = pika_treatment, y = cover, color = posion_plant_treatment)) +
    geom_jitter(
      size = 1, alpha = 0.4,
      position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7)
    ) +
    geom_point(
      data = letters_df,
      aes(x = pika_treatment, y = response * 100, fill = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 8, pch = 21, color = "black"
    ) +
    geom_text(
      data = letters_df,
      aes(x = pika_treatment, y = text_y, label = .group, group = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 5, color = "black", show.legend = FALSE
    ) +
    scale_fill_manual(values = c(green, pink)) +
    scale_color_manual(values = c(green, pink)) +
    labs(
      y = y_label,
      x = "",
      color = "Poison plant treatment",
      fill = "Poison plant treatment"
    ) +
    theme_pubr(base_size = 15) +
    theme(legend.position = "bottom", legend.title = element_blank())
}

# Now call the function with appropriate arguments
forb_plot    <- make_cover_plot("forbs", forbs_letters, "forb cover (%)", text_y = 41)
grass_plot   <- make_cover_plot("grasses", grasses_letters, "grass cover (%)", text_y = 41)
poison_plant <- make_cover_plot("s_chamaejasme", poison_letters, "S. chamaejasme cover (%)", text_y = 43)
sedge_plot   <- make_cover_plot("sedges", sedges_letters, "sedge cover (%)", text_y = 70)



(forb_plot + grass_plot) / (poison_plant + sedge_plot) +
    plot_annotation(tag_levels = 'A') + 
    plot_layout(guides = "collect") & theme(legend.position = 'bottom')



```


### Forbs Cover

There was a significant negative effect of *S. chamaejasme* treatment on forb cover, and a marginally significant positive interaction between pika presence and *S. chamaejasme* treatment (Table @tbl-forbs-model-summary). Post-hoc contrasts showed increased forb cover in the *S. chamaejasme* treatment, but this was modulated by pika presence (Table @tbl-forbs-model-contrasts).


```{r}
#| label: tbl-forbs-model-summary
#| tbl-cap: "Model summary for forb cover."
#| echo: false
#| message: false
#| warning: false

library(kableExtra)
forbs_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-forbs-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for forb cover by treatment group."
#| echo: false
#| message: false
#| warning: false

forbs_model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


### Grasses Cover

Grasses responded strongly to the *S. chamaejasme* treatment, showing a large increase in cover under *S. chamaejasme* alone and a significant negative interaction with pika presence (Table @tbl-grasses-model-summary). Pairwise contrasts confirm that the strongest grass suppression occurred when both pikas and *S. chamaejasme* were present (Table @tbl-grasses-model-contrasts).


```{r}
#| label: tbl-grasses-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false

grasses_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-grasses-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

grasses_model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


### *S. chamaejasme* Cover

There was a significant **negative effect** of *S. chamaejasme* treatment, but pika presence had no clear effect (Table @tbl-poison-model-summary). No post-hoc contrasts were statistically significant (Table @tbl-poison-model-contrasts).



```{r}
#| label: tbl-poison-model-summary
#| tbl-cap: "Model summary for *S. chamaejasme* cover."
#| echo: false
#| message: false
#| warning: false

poison_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-poison-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for *S. chamaejasme* cover by treatment group."
#| echo: false
#| message: false
#| warning: false

poison__model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


### Sedges Cover

Sedge cover declined significantly under *S. chamaejasme* treatment butincreased when pikas were also present, indicating a strong positive interaction effect (Table @tbl-sedges-model-summary). Post-hoc contrasts confirm the synergistic effect of pika presence and *S. chamaejasme* on sedge abundance (Table @tbl-sedges-model-contrasts).


```{r}
#| label: tbl-sedges-model-summary
#| tbl-cap: "Model summary for sedge cover."
#| echo: false
#| message: false
#| warning: false

sedges_model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-sedges-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for sedge cover by treatment group."
#| echo: false
#| message: false
#| warning: false

sedges_model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


## Experiment: Foraging Efficiency

### Bite Steps

```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

# Load the bite_steps dataset once
bite_steps <- read_csv(here("data/processed/experiment_foraging_efficiency/yak_bite_steps_ratio.csv"))

# Define a function to load model data for a plant group
load_bite_step_data <- function(plant) {
  base_path <- here(glue::glue("data/processed/experiment_foraging_efficiency/modeled_data/bite_steps/{plant}_bite_steps"))
  list(
    emms = read_csv(file.path(base_path, "emms.csv")),
    letters = read_csv(file.path(base_path, "posthoc_letters.csv")),
    model_summary = read_csv(file.path(base_path, "model_summary.csv")),
    model_contrasts = read_csv(file.path(base_path, "posthoc_contrasts.csv"))
  )
}

# Define your plant groups
plant_groups <- c("forbs", "grasses", "sedges")

# Load all model data into a named list
bite_step_models <- map(plant_groups, load_bite_step_data) |> set_names(plant_groups)

```



The following figure (@fig-bite_step_figures) shows the relationship between yak bite/step ratios per plant across he pika and poison plant treatments. We see decreased bite/step ratios for grasses and sedges within the no pika and poison plant treatment plots. Whereas all other treatment levels had a consistent bite/step ratio. Forbs is an exception where there is no difference except with increase forb bite/step ratio in the pika x poison plant treatment level.


```{r}
#| label: fig-bite_step_figures
#| fig-cap: "diet selection"
#| echo: false
#| reulsts: asis'
#| message: false
#| warning: false

# Define your color palette
plant_colors <- c(green, pink)

# Reusable function
plot_bite_steps <- function(plant_name, y_label, label_y = NULL) {
  data <- bite_steps |> filter(plant == plant_name)
  emms <- bite_step_models[[plant_name]]$emms
  letters <- bite_step_models[[plant_name]]$letters

  label_y <- label_y %||% max(data$bites_steps_ratio, na.rm = TRUE) + 0.05

  data |>
    left_join(emms, by = c('pika_treatment', 'posion_plant_treatment')) |>
    ggplot(aes(x = pika_treatment, y = bites_steps_ratio, color = posion_plant_treatment)) +
    geom_jitter(
      size = 1, alpha = 0.4,
      position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7)
    ) +
    geom_point(
      data = letters,
      aes(x = pika_treatment, y = response, fill = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 8, pch = 21, color = "black", inherit.aes = FALSE
    ) +
    geom_text(
      data = letters,
      aes(x = pika_treatment, y = label_y, label = .group, group = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 5, color = "black", show.legend = FALSE, inherit.aes = FALSE
    ) +
    scale_fill_manual(values = plant_colors) +
    scale_color_manual(values = plant_colors) +
    labs(
      y = y_label,
      x = "",
      color = "Poison plant treatment",
      fill = "Poison plant treatment"
    ) +
    theme_pubr(base_size = 15) +
    theme(legend.position = "bottom", legend.title = element_blank())
}

# Call the function for each plant type
grasses_plot <- plot_bite_steps("grasses", "grass bite/steps")
sedges_plot  <- plot_bite_steps("sedges", "sedges bite/steps", label_y = 1)
forbs_plot   <- plot_bite_steps("forbs", "forbs bite/steps", label_y = 0.4)



(grasses_plot + sedges_plot) / (forbs_plot + plot_spacer()) + 
  plot_annotation(tag_levels = 'A') + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')



```


#### grass bite steps


The model revealed that the presence of S. chamaejasme significantly reduced yak grass feeding efficiency (bite/step ratio) (@tbl-grasses-bite-step-model-summary), but this negative effect was mitigated when pikas were present, as shown by a significant positive interaction (@tbl-grasses-bite-step-model-summary). Pika presence alone did not influence feeding efficiency, but in combination with S. chamaejasme, it substantially improved it (@tbl-grasses-bite-step-model-contrasts). This suggests that pikas may facilitate yak foraging by offsetting the detrimental effects of poisonous plants.


```{r}
#| label: tbl-grasses-bite-step-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false

bite_step_models$grasses$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-grasses-bite-step-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

bite_step_models$grasses$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


#### sedge bite steps

The presence of S. chamaejasme significantly reduced yak feeding efficiency on sedges, as shown by a strong negative main effect in the model summary (see @tbl-sedges-bite-step-model-summary). Although pika presence alone had no effect, the significant positive interaction indicates that pikas mitigated the negative impact of the poisonous plant. Specifically, sedge bite/step ratios were much lower when S. chamaejasme was present without pikas, but this reduction was substantially less when pikas were present (see pairwise contrasts in @tbl-sedges-bite-step-model-contrasts). These findings suggest that pikas facilitate yak foraging on sedges by reducing the negative effects of *S. chamaejasme*.

```{r}
#| label: tbl-sedges-bite-step-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false

bite_step_models$sedges$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-sedges-bite-step-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

bite_step_models$sedges$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


#### forbs bite steps

For forbs, neither pika presence nor S. chamaejasme alone significantly affected yak feeding efficiency, as both main effects were non-significant (see @tbl-forbs-bite-step-model-summary). However, there was a marginally non-significant interaction (p = 0.067), suggesting a possible moderating effect of pikas on the impact of the poisonous plant. In the pairwise contrasts, the bite/step ratio was significantly lower when S. chamaejasme was present and pikas were absent compared to when both were present (see @tbl-forbs-bite-step-model-contrasts). This indicates a weaker but still detectable facilitative effect of pikas on yak forb consumption in the presence of S. chamaejasme.

```{r}
#| label: tbl-forbs-bite-step-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false

bite_step_models$forbs$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-forbs-bite-step-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

bite_step_models$forbs$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

### Plant bites


```{r}
#| echo: false
#| results: "hide"
#| message: false
#| warning: false

# Load the bite_steps dataset once
plant_bites <- read_csv(here("data/processed/experiment_foraging_efficiency/yak_plant_bites.csv")) |>
  mutate(plant = gsub("_[^_]*$", '',plant))

# Define a function to load model data for a plant group
load_plant_bites_data <- function(plant) {
  base_path <- here(glue::glue("data/processed/experiment_foraging_efficiency/modeled_data/plant_bites/{plant}_plant_bites"))
  list(
    emms = read_csv(file.path(base_path, "emms.csv")),
    letters = read_csv(file.path(base_path, "posthoc_letters.csv")),
    model_summary = read_csv(file.path(base_path, "model_summary.csv")),
    model_contrasts = read_csv(file.path(base_path, "posthoc_contrasts.csv"))
  )
}

# Define your plant groups
plant_groups <- c("forbs", "grasses", "sedges")

# Load all model data into a named list
plant_bites_models <- map(plant_groups, load_plant_bites_data) |> set_names(plant_groups)

```



The following figure (@fig-bite_step_figures) shows the relationship between yak bite/step ratios per plant across he pika and poison plant treatments. We see decreased bite/step ratios for grasses and sedges within the no pika and poison plant treatment plots. Whereas all other treatment levels had a consistent bite/step ratio. Forbs is an exception where there is no difference except with increase forb bite/step ratio in the pika x poison plant treatment level.


```{r}
#| label: fig-plant_bites_figures
#| fig-cap: "yak plant bites by plant group"
#| echo: false
#| reulsts: asis
#| message: false
#| warning: false

# Define your color palette
plant_colors <- c(green, pink)

plant_bites_models$forbs$emms
# Reusable function

plot_plant_bites <- function(plant_name, y_label, label_y = NULL) {
  data <- plant_bites |> filter(plant == plant_name)
  emms <- plant_bites_models[[plant_name]]$emms
  letters <- plant_bites_models[[plant_name]]$letters

  if ("emmean" %in% names(emms)) {
    emms <- emms |> rename(response = emmean)
    letters <- letters |> rename(response = emmean)
  }

  label_y <- label_y %||% max(data$forage_efficiency, na.rm = TRUE) + 0.05

  data |>
    left_join(emms, by = c('pika_treatment', 'posion_plant_treatment')) |>
    ggplot(aes(x = pika_treatment, y = forage_efficiency, color = posion_plant_treatment)) +
    geom_jitter(
      size = 1, alpha = 0.4,
      position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.7)
    ) +
    geom_point(
      data = letters,
      aes(x = pika_treatment, y = response, fill = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 8, pch = 21, color = "black", inherit.aes = FALSE
    ) +
    geom_text(
      data = letters,
      aes(x = pika_treatment, y = label_y, label = .group, group = posion_plant_treatment),
      position = position_dodge(width = 0.7),
      size = 5, color = "black", show.legend = FALSE, inherit.aes = FALSE
    ) +
    scale_fill_manual(values = plant_colors) +
    scale_color_manual(values = plant_colors) +
    labs(
      y = y_label,
      x = "",
      color = "Poison plant treatment",
      fill = "Poison plant treatment"
    ) +
    theme_pubr(base_size = 15) +
    theme(legend.position = "bottom", legend.title = element_blank())
}

# Call the function for each plant type
grasses_plot <- plot_plant_bites("grasses", "grass bites", label_y = 350)
sedges_plot  <- plot_plant_bites("sedges", "sedge bites", label_y = 410)
forbs_plot   <- plot_plant_bites("forbs", "forb bites", label_y = 200)



(grasses_plot + sedges_plot) / (forbs_plot + plot_spacer()) + 
  plot_annotation(tag_levels = 'A') + 
  plot_layout(guides = "collect") & theme(legend.position = 'bottom')



```




#### grass plant bites


Yak grazing on grasses was significantly reduced by the presence of the poisonous plant S. chamaejasme, with a large negative estimate (-45.3) indicating fewer bites in treated plots. The pika treatment alone had no significant effect on grass bites. However, the interaction between pika presence and S. chamaejasme treatment was positive and significant (+43.9), suggesting that pika presence mitigates the negative effect of the poisonous plant on grass bites. Contrasts confirm that plots with S. chamaejasme had substantially lower grazing, except when pika were present, which restored grazing levels.

```{r}
#| label: tbl-grasses-plant-bites-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false


plant_bites_models$grasses$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-grasses-plant-bite-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

plant_bites_models$grasses$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


#### Forb plant bites

Yak bites on forbs were generally low and showed no significant differences across pika or poisonous plant treatments. The S. chamaejasme treatment showed a marginally positive effect on forb bites, but this was not statistically significant (p ~ 0.058). The interaction between pika and S. chamaejasme was also non-significant. Overall, yak grazing on forbs appears unaffected by either treatment or their interaction.

```{r}
#| label: tbl-forbs-plant-bites-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false


plant_bites_models$forbs$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-forbs-plant-bite-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

plant_bites_models$forbs$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```


#### Sedge plant bites

Grazing on sedges was significantly reduced by S. chamaejasme treatment (estimate = -0.42), indicating fewer yak bites in plots with the poisonous plant. Pika presence alone had no significant effect, but the pika × S. chamaejasme interaction was strongly positive (+0.36), showing that pika presence offsets the reduction caused by the poisonous plant. Contrasts reveal that sedge grazing decreases with S. chamaejasme unless pika are present, which helps maintain grazing levels.


```{r}
#| label: tbl-sedges-plant-bites-model-summary
#| tbl-cap: "Model summary for grass cover."
#| echo: false
#| message: false
#| warning: false


plant_bites_models$sedges$model_summary |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```

```{r}
#| label: tbl-sedges-plant-bite-model-contrasts
#| tbl-cap: "Estimated marginal means contrasts for grass cover by treatment group."
#| echo: false
#| message: false
#| warning: false

plant_bites_models$sedges$model_contrasts |>
  kable(format = "latex", booktabs = TRUE, digits = 3) |>
  kable_styling(latex_options = c("striped"))
```
